\subsection{Ordinary Differential Equations}
A \textbf{differential equation} is an equation that relates some unknown
functions and their derivatives.
An \textbf{ordinary differential equation} (ODE) is a differential equation
that relates one or more functions of a single variable and their ordinary
derivatives.
The \textbf{order} of a differential equation is the highest order of the
derivative in the equation.
Newton's second law of motion, the force \(F\) acting on an object with mass
\(m\) and acceleration \(a\) is \(F=ma\) can be written as an ordinary
differential equation
\[F=m\frac{dv}{dt}\]
which is a first order differential equation, or
\[F=m\frac{d^2s}{dt^2}\]
 which is a second order differential equation.
By Newton's second law of motion, the net force \(F\) acting on a falling
object with mass \(m\), velocity \(v\), and air resistance force \(\gamma v\)
can be modeled by the differential equation
\[F=m\frac{dv}{dt}=mg-\gamma v\]
where \(g\) is the acceleration due to gravity and \(\gamma\) is the drag
coefficient.
A function \(f\) is a \textbf{solution} of a differential equation if the
function and its derivatives satisfy the equation for all values of \(x\) in
some open interval \(a<x<b\).
It is possible that there are many solutions of a differential equation.
An initial condition is a condition \(y(x_0)=y_0\) or
\(y^{(n)}(x_0)=y_n\) on the solution.
An \textbf{initial value problem} is solving a differential equation with initial conditions.
The interval of validity is the largest possible interval on which
the solution is valid and contains \(x_0\) in the initial conditions.
The general solution of a differential equation is the set of all solutions
and the particular solution is the solution that satisfies the initial
conditions.
An explicit solution is any solution in the form \(y=y(x)\), otherwise it is
an implicit solution.
The particular solution of the differential equation
\[\frac{dy}{dx}=y\]
with initial
condition \(y(0)=1\) is \(y=e^x\) since
\[\frac{dy}{dx}=\frac{d}{dx}e^x=e^x=y\]
and \(y(0)=e^0=1\).
The \textbf{existence} and \textbf{uniqueness} problem asks that given a
differential equation, does there exist a solution and if any is there only
one solution.