\subsection{Separable Equations}

\subsubsection*{Separable Equations}
A \textbf{separable equation} is a first order differential equation that can
be written in the differential form
\[M(x)\,dx+N(y)\,dy=0\]
or can be written in the form
\[\frac{dy}{dx}=f(x)g(y)\]
We can separate the variables if \(g(y)\neq0\), let \(h(y)=1/g(y)\) then
\[\frac{dy}{dx}=\frac{f(x)}{h(y)}\]
We write the equation in the differential form
\[h(y)\,dy=f(x)\,dx\]
Then we integrate both sides of the equation:
\[\int h(y)\,dy=\int f(x)\,dx\]
Now we have an implicit solution of the differential equation and sometimes we
can solve for an explicit solution.
We can justify the method of separation of variables by using the chain rule
to show that
\begin{align*}
    \frac{d}{dx}\int h(y)\,dy &= \frac{d}{dx}\int f(x)\,dx \\
    \frac{d}{dy}\int h(y)\,dy\frac{dy}{dx} &= f(x) \\
    h(y)\frac{dy}{dx} &= f(x) \\
    \frac{dy}{dx} &= \frac{f(x)}{h(y)}=f(x)g(y)
\end{align*}
\begin{problem}
    Solve the differential equation \(\dfrac{dy}{dx}=-xy\).
\end{problem}
\begin{solution}
    Notice that \(y=0\) is a trivial solution then we solve the differential
    equation for non-trivial solutions \(y\neq 0\).
    We use separation of variables then
    \begin{align*}
        \dfrac{dy}{dx} &= -xy \\
        \int\frac{dy}{y} &= -\int x\,dx \\
        \ln |y|+C_1 &= -\frac{x^2}{2}+C_2
    \end{align*}
    Let \(C=C_2-C_1\), then
    \begin{align*}
        \ln |y| &= -\frac{x^2}{2}+C_2-C_1=-\frac{x^2}{2}+C \\
        |y| &= e^{(-x^2/2)+C}=e^Ce^{-x^2/2} \\
        y &= \pm e^Ce^{-x^2/2}=Ae^{-x^2/2}
    \end{align*}
    where \(A\in\R\) is an arbitrary constant.
\end{solution}
\begin{problem}
    Solve the differential equation \(\dfrac{dy}{dx}=\dfrac{x^2}{y^2}\) with
    the initial condition \(y(0)=2\).
\end{problem}
\begin{solution}
    We use separation of variables to find the general solution then
    \begin{align*}
        \dfrac{dy}{dx} &= \dfrac{x^2}{y^2} \\
        \int y^2\,dy &= \int x^2\,dx \\
        \frac{y^3}{3}+C_1 &= \frac{x^3}{3}+C_2 \\
        \frac{y^3}{3} &=\frac{x^3}{3}+C,\quad C=C_2-C_1 \\
        y^3 &= x^3+3C \\
        y &= \sqrt[3]{x^3+K},\quad K=3C
    \end{align*}
    We consider the initial condition \(y(0)=2\) to find the particular
    solution then
    \[y(0)=\sqrt[3]{0+K}\iff 2=\sqrt[3]{K}\iff K=8\]
    The solution of the initial value problem is \(y=\sqrt[3]{x^3+8}\).
\end{solution}
\begin{problem}
    Solve the differential equation \(\dfrac{dy}{dx}=\dfrac{6x^2}{2y+\cos y}\).
\end{problem}
\begin{solution}
    We have
    \begin{align*}
        \dfrac{dy}{dx} &= \dfrac{6x^2}{2y+\cos y} \\
        \int(2y+\cos y)\,dy &= \int 6x^2\,dx \\
        y^2+\sin y &= 2x^3+C
    \end{align*}
    where \(C\in\R\) is an arbitrary constant.
\end{solution}
\begin{problem}
    Solve the differential equation \(y'=x^2y\).
\end{problem}
\begin{solution}
    We have
    \begin{align*}
        \frac{dy}{dx} &= x^2y \\ \int \frac{dy}{y} &= \int x^2\,dx \\
        \ln |y| &= \frac{x^3}{3}+C \\ |y| &= e^{(x^3/3)+C}\\ y &= Ae^{x^3/3}
    \end{align*}
    where \(A\in\R\) is an arbitrary constant.
\end{solution}

\subsubsection*{Homogeneous Equations}
A \textbf{homogeneous} equation is in the form
\[\frac{dy}{dx}=f\left(\frac{y}{x}\right)\]
We can transform a homogeneous equation into a separable equation by a change
of variable.
Consider the equation
\[\frac{dy}{dx}=\frac{y-4x}{x-y}\]
and we can show that
\[\frac{dy}{dx}=\frac{(y/x)-4}{1-(y/x)}\]
thus the equation is homogeneous.
Let \(v=y/x\iff y=vx\) so
\begin{align*}
    \frac{dy}{dx}=\frac{dv}{dx}x+v &= \frac{v-4}{1-v} \\
    \frac{dv}{dx}x
    &= \frac{v-4}{1-v}-v=\frac{v-4-v(1-v)}{1-v}=\frac{v^2-4}{1-v}
\end{align*}
and the equation is separable then
\begin{align*}
    \int\frac{1-v}{v^2-4}\,dv &= \int\frac{dx}{x}
\end{align*}
Since
\[\int\frac{1-v}{v^2-4}\,dv=\int\frac{dv}{v^2-4}-\int\frac{v}{v^2-4}\,dv\]
then
\begin{align*}
    \frac{1}{v^2-4}=\frac{1}{(v+2)(v-2)} &= \frac{A}{v+2}+\frac{B}{v-2} \\
    1 &= A(v-2)+B(v+2) \\
    v=-2\iff -4A=1\iff A &= -\frac{1}{4} \\
    v=2\iff 4B=1\iff B &= \frac{1}{4}
\end{align*}
so
\[\int\frac{dv}{v^2-4}
=-\frac{1}{4}\int\frac{dv}{v+2}+\frac{1}{4}\int\frac{dv}{v-2}
=-\frac{1}{4}\ln|v+2|+\frac{1}{4}\ln|v-2|\]
and
\[\int\frac{v}{v^2-4}\,dv=\frac{1}{2}\ln|v^2-4|\]
therefore
\[\int\frac{1-v}{v^2-4}\,dv
=-\frac{1}{4}\ln|v+2|+\frac{1}{4}\ln|v-2|-\frac{1}{2}\ln|v^2-4|\]
Then
\begin{align*}
    -\frac{1}{4}\ln|v+2|+\frac{1}{4}\ln|v-2|-\frac{1}{2}\ln|v^2-4|
    &= \ln|x|+C_1 \\
    \ln|v-2|-\ln|v+2|-2\ln|v^2-4| &= 4\ln|x|+C_2,\quad C_2=4C_1 \\
    \ln\left|\frac{v-2}{v+2}\right|-\ln((v^2-4)^2)
    =\ln\left|\frac{v-2}{(v+2)(v+2)^2(v-2)^2}\right| &= \ln(x^4)+C_2 \\
    \ln\left|\frac{1}{(v+2)^3(v-2)}\right| &= \ln(x^4)+C_2 \\
    \frac{1}{(v+2)^3(v-2)} &= C_3x^4,\quad C_3=\pm e^{C_2} \\
    (v+2)^3(v-2)x^4 &= C,\quad C=1/C_3 \\
    (vx+2x)^3(vx-2x) &= C
\end{align*}
Therefore the solution is
\[(y+2x)^3(y-2x)=C\]
\begin{problem}
    Solve the differential equation
    \[\frac{dy}{dx}=\frac{x^2+xy+y^2}{x^2}\]
\end{problem}
\begin{solution}
    Since
    \[\frac{dy}{dx}=\frac{x^2+xy+y^2}{x^2}
    =1+\frac{y}{x}+\left(\frac{y}{x}\right)^2\]
    hence the equation is homogeneous.
    Let \(y=vx\), then
    \begin{align*}
        \frac{dy}{dx}=\frac{dv}{dx}x+v &= 1+v+v^2 \\
        \frac{dv}{dx}x &= 1+v^2
    \end{align*}
    and by separation of variables
    \begin{align*}
        \int\frac{dv}{1+v^2} &= \int\frac{dx}{x} \\
        \arctan v &= \ln x+C \\
        \arctan\left(\frac{y}{x}\right)-\ln x &= C \qedhere
    \end{align*}
\end{solution}

\subsubsection*{Orthogonal Trajectories}
An \textbf{orthogonal trajectory} of a family of curves is a curve that
intersects each curve of the family orthogonally, that is, at right angles.
\begin{problem}
    Find the orthogonal trajectories of the family of curves \(x=ky^2\), where
    \(k\) is an arbitrary constant.
\end{problem}
\begin{solution}
    The curves \(x=ky^2\) form a family of parabolas whose axis of symmetry is
    the \(x\)-axis.
    We differentiate \(x=ky^2\) and have
    \begin{align*}
        1 &= 2ky\frac{dy}{dx} \\
        \frac{dy}{dx} &= \frac{1}{2ky}
    \end{align*}
    and since \(k=x/y^2\) so we get
    \[\frac{dy}{dx}=\frac{1}{2(x/y^2)y}=\frac{y}{2x}\]
    which is the slope of the tangent line at any point \((x,y)\) on one of
    the parabolas.
    On an orthogonal trajectory the slope of the tangent line must be the
    negative reciprocal of this slope.
    Therefore the orthogonal trajectories must satisfy the differential
    equation
    \[\frac{dy}{dx}=-\frac{2x}{y}\]
    This differential equation is separable, and we solve it as follows:
    \begin{align*}
        \int y\,dy &= -\int 2x\,dx \\
        \frac{y^2}{2} &= -x^2+C \\
        x^2+\frac{y^2}{2} &= C
    \end{align*}
    where \(C\) is an arbitrary positive constant.
    Thus the orthogonal trajectories are the family of ellipses given by the
    equation \(x^2+(y^2/2)=C\).
\end{solution}

\subsubsection*{Mixing Problems}