\subsection{Partial Fractions}
\subsubsection{Partial Fractions}
Consider a rational function \(\displaystyle{f(x)=\frac{P(x)}{Q(x)}}\) where
\(P\) and \(Q\) are polynomials.
If \(f\) is improper, that is, \(\deg(P)\geq\deg(Q)\), then we simplify to get
\(\displaystyle{f(x)=S(x)+\frac{R(x)}{Q(x)}}\) where \(S\) and \(R\) are
polynomials.
Then we can factor \(Q\) to be irreducible and express \(\dfrac{R(x)}{Q(x)}\)
as the a sum of partial fractions of the form
\[\frac{A}{(ax+b)^i}\quad\text{or}\quad\frac{Ax+B}{(ax^2+bx+c)^j}\]
There are four possible cases:
\begin{enumerate}
    \item \(Q\) is a product of distinct linear factors.
    Then
    \[Q(x)=(a_1 x+b_1)(a_2 x+b_2)\cdots(a_k x+b_k)\]
    and therefore
    \[\frac{R(x)}{Q(x)}=\frac{A_1}{a_1 x+b_1}+\frac{A_2}{a_2 x+b_2}+\cdots+
    \frac{A_k}{a_k x+b_k}\]
    \item \(Q\) is a product of linear factors where some are repeated.
    Suppose that the first linear factor is repeated \(r\) times, then we have
    \[\frac{A_1}{a_1 x+b_1}+\frac{A_2}{(a_1 x+b_1)^2}+\cdots+
    \frac{A_r}{(a_1 x+b_1)^r}\]
    for the first repeated linear factor and similarly for other repeated
    linear factors.
    \item \(Q\) has irreducible quadratic factors without repeated
    factors.
    Then for every quadratic factor we have
    \[\frac{Ax+B}{ax^2+bx+c}\]
    \item \(Q\) has a repeated irreducible quadratic factor.
    Suppose that the first quadratic factor is repeated \(r\) times,
    then we have
    \[\frac{A_1 x+B_1}{ax^2+bx+c}+\frac{A_2 x+B_2}{(ax^2+bx+c)^2}+\cdots
    +\frac{A_r x+B_r}{(ax^2+bx+c)^r}\]
    for the first repeated quadratic factor and similarly for the others.
\end{enumerate}
\begin{problem}
    Evaluate \(\displaystyle{\int\frac{x^2+2x-1}{2x^3+3x^2-2x}\,dx}\).
\end{problem}
\begin{solution}
    We simplify to get \(2x^3+3x^2-2x=x(2x-1)(x+2)\) then
    \begin{align*}
        \frac{x^2+2x-1}{x(2x-1)(x+2)} &=
        \frac{A}{x}+\frac{B}{2x-1}+\frac{C}{x+2} \\
        x^2+2x-1 &= A(2x-1)(x+2)+Bx(x+2)+Cx(2x-1) \\
        x=0\iff -1=-2A &\iff A=\frac{1}{2} \\
        x=\frac{1}{2}\iff \frac{1}{4}=\frac{5}{4}B &\iff B=\frac{1}{5} \\
        x=-2\iff -1=10C &\iff C=-\frac{1}{10}
    \end{align*}
    and therefore
    \begin{align*}
        \int\frac{x^2+2x-1}{2x^3+3x^2-2x}
        &= \frac{1}{2}\int\frac{dx}{x}+\frac{1}{5}\int\frac{dx}{2x-1}
        -\frac{1}{10}\int\frac{dx}{x+2} \\
        &= \frac{1}{2}\ln|x|+\frac{1}{10}\ln|2x-1|
        -\frac{1}{10}\ln|x+2|+K \qedhere
    \end{align*}
\end{solution}
\begin{problem}
    Evaluate \(\displaystyle{\int\frac{x^4-2x^2+4x+1}{x^3-x^2-x+1}\,dx}\).
\end{problem}
\begin{solution}
    We simplify to get
    \begin{align*}
        x^4-2x^2+4x+1 &=(x^2-1)^2+4x=(x+1)^2(x-1)^2+4x \\
        x^3-x^2-x+1 &= x^2(x-1)-(x-1)=(x^2-1)(x-1)=(x+1)(x-1)^2
    \end{align*}
    and so
    \begin{align*}
        \int\frac{x^4-2x^2+4x+1}{x^3-x^2-x+1}\,dx
        &= \int(x+1)\,dx+\int\frac{4x}{(x+1)(x-1)^2}\,dx
    \end{align*}
    Then
    \begin{align*}
        \frac{4x}{(x+1)(x-1)^2} &=
        \frac{A}{x+1}+\frac{B}{x-1}+\frac{C}{(x-1)^2} \\
        4x &= A(x-1)^2+B(x+1)(x-1)+C(x+1) \\
        x=-1\iff -4=4A &\iff A=-1 \\ x=1\iff 4=2C &\iff C=2 \\
        x=0\iff 0=-1-B+2 &\iff B=1
    \end{align*}
    and therefore
    \begin{align*}
        \int\frac{x^4-2x^2+4x+1}{x^3-x^2-x+1}\,dx
        &= \int(x+1)\,dx-\int\frac{dx}{x+1}+\int\frac{dx}{x-1}
        +2\int\frac{dx}{(x-1)^2} \\
        &= \frac{x^2}{2}+x-\ln|x+1|+\ln|x-1|-\frac{2}{x-1}+K \qedhere
    \end{align*}
\end{solution}
\begin{problem}
    Evaluate \(\displaystyle{\int\frac{2x^2-x+4}{x^3+4x}\,dx}\).
\end{problem}
\begin{solution}
    We do partial fraction decomposition then
    \begin{align*}
        \frac{2x^2-x+4}{x^3+4x} &= \frac{2x^2-x+4}{x(x^2+4)}
        =\frac{A}{x}+\frac{Bx+C}{x^2+4} \\
        2x^2-x+4 &= A(x^2+4)+(Bx+C)x=(A+B)x^2+Cx+4A \\
        A &=1,\quad B=1,\quad C=-1
    \end{align*}
    and so
    \begin{align*}
        \int\frac{2x^2-x+4}{x^3+4x}\,dx
        &= \int\frac{dx}{x}+\int\frac{x-1}{x^2+4}\,dx
        =\ln|x|+\frac{1}{2}\ln|x^2+4|
        +\frac{1}{2}\arctan\left(\frac{x}{2}\right)+K \qedhere
    \end{align*}
\end{solution}
\begin{problem}
    Evaluate \(\displaystyle{\int\frac{4x^2-3x+2}{4x^2-4x+3}\,dx}\).
\end{problem}
\begin{solution}
    
\end{solution}
\begin{problem}
    Evaluate \(\displaystyle{\int\frac{1-x+2x^2-x^3}{x(x^2+1)^2}\,dx}\).
\end{problem}
\begin{solution}
    We do partial fraction decomposition then
    \begin{align*}
        \frac{1-x+2x^2-x^3}{x(x^2+1)^2}
        &= \frac{A}{x}+\frac{Bx+C}{x^2+1}+\frac{Dx+E}{(x^2+1)^2} \\
        1-x+2x^2-x^3 &= A(x^2+1)^2+(Bx+C)x(x^2+1)+(Dx+E)x \\
        x=0\iff A&=1
    \end{align*}
    and so
    \begin{align*}
        1-x+2x^2-x^3 &=
    \end{align*}
\end{solution}