\subsection{Derivatives}
\subsubsection{Derivatives and Rates of Change}
\begin{definition}
    The \textbf{tangent line} of the curve \(y=f(x)\) at the point \(P(a, f(a))\)
    is the line through \(P\) with slope
    \[m=\lim_{x\to a}\frac{f(x)-f(a)}{x-a}=\lim_{h\to 0}\frac{f(a+h)-f(a)}{h}\]
    where \(h=x-a\), if this limit exists.
\end{definition}
\begin{definition}
    The \textbf{velocity} at time \(t=a\) of a \textbf{position function}
    \(s=f(t)\) is \[v(a)=\lim_{h\to 0}\frac{f(a+h)-f(a)}{h}\qedhere\]
\end{definition}
\begin{definition}
    The \textbf{derivative} of a function \(f\) at a number \(a\) is
    \[f'(a)=\lim_{h\to 0}\frac{f(a+h)-f(a)}{h}\] if this limit exists.
\end{definition}
The slope of the tangent line to \(y=f(x)\) at the point \((a,f(a))\) is
\(f'(a)\), the derivative of \(f\) at \(a\).
The equation of the tangent line is \[y-f(a)=f'(a)(x-a)\]
The \textbf{instantaneous rate of change} of \(y=f(x)\) with respect to \(x\)
at \(x=x_0\) in the interval \([x_0,x_1]\) is
\[\lim_{\Delta x\to 0}\frac{\Delta y}{\Delta x}
=\lim_{x_1 \to x_0}\frac{f(x_1)-f(x_0)}{x_1-x_0}\]
The derivative \(f'(a)\) is the instantaneous rate of change of \(y=f(x)\)
with respect to \(x\) when \(x=a\).
If \(s=f(t)\) is a position function of an object, then the velocity of the
object at time \(t=a\) is \(v(a)=f'(a)\) and the \textbf{speed} of the
object is \(|f'(a)|\), the \textbf{magnitude} of the velocity.

\subsubsection{The Derivative as a Function}
\begin{definition}
    The derivative of a function \(f(x)\) is the function
    \[f'(x)=\lim_{h\to 0}\frac{f(x+h)-f(x)}{h}\qedhere\]
\end{definition}
The following notations of the derivative of \(y=f(x)\) with respect to \(x\) are
equivalent.
\[y'=f'(x)=\frac{dy}{dx}=\frac{d}{dx}f(x)=D_x f(x)\]
where \(f'(x)\) is Newton's notation and \(dy/dx\) is Leibniz's notation.
The notation \(d/dx\) is the \textbf{differential operator} that indicates the
operation of \textbf{differentiation}.
The notations of the derivative of \(f(x)\) at \(a\) are
\[f'(a)=\left.\frac{dy}{dx}\right|_{x=a}=\left[\frac{dy}{dx}\right]_{x=a}\]
\begin{problem}
    Find the derivative of \(f(x)=\sqrt{x}\).
\end{problem}
\begin{solution}
    We have
    \begin{align*}
        f'(x)&=\lim_{h\to 0}\frac{\sqrt{x+h}-\sqrt{x}}{h}=\lim_{h\to 0}
        \left(\frac{\sqrt{x+h}-\sqrt{x}}{h}\cdot
        \frac{\sqrt{x+h}+\sqrt{x}}{\sqrt{x+h}+\sqrt{x}}\right) \\
        &=\lim_{h\to 0}\frac{x+h-x}{h(\sqrt{x+h}+\sqrt{x})}
        =\lim_{h\to 0}\frac{1}{\sqrt{x+h}+\sqrt{x}}=\frac{1}{\sqrt{x}+\sqrt{x}}
        =\frac{1}{2\sqrt{x}}
    \end{align*}
    The derivative of \(f(x)=\sqrt{x}\) is \[f'(x)=\frac{1}{2\sqrt{x}}\qedhere\]
\end{solution}

\subsubsection{Differentiable Functions}
\begin{definition}
    A function \(f\) is \textbf{differentiable} at \(a\) if \(f'(a)\) exists.
    \(f\) is differentiable on an open interval if it is differentiable at
    every number in the interval.
\end{definition}
\begin{theorem}
    If \(f\) is differentiable at \(a\), then \(f\) is continuous at \(a\).
\end{theorem}
\begin{proof}
    Given that \(f\) is differentiable at \(a\), we want to show that
    \[\lim_{x\to a}f(x)=f(a)\]
    Since \(f'(a)\) exists, we have
    \[f'(a)=\lim_{x\to a}\frac{f(x)-f(a)}{x-a}\]
    Then we have
    \[\lim_{x\to a}[f(x)-f(a)]=\lim_{x\to a}\left(\frac{f(x)-f(a)}{x-a}(x-a)\right)
    =\lim_{x\to a}\left(\frac{f(x)-f(a)}{x-a}\right)\lim_{x\to a}(x-a)
    =f'(a)\cdot 0=0\]
    Then we have
    \[\lim_{x\to a}f(x)=\lim_{x\to a}(f(a)+f(x)-f(a))
    =\lim_{x\to a}f(a)+\lim_{x\to a}(f(x)-f(a))=f(a)+0=f(a)\]
    Therefore, it is proved that \(f\) is continuous at \(a\).
\end{proof}
Note that there are functions that are continuous but not differentiable.
The function \(y=|x|\) is continuous at 0 but not differentiable at 0 since
\[f'(0)=\lim_{h\to 0}\frac{|0+h|-|0|}{h}\] if the limit exists but
\begin{align*}
    \lim_{h\to 0^-}\frac{|0+h|-|0|}{h}&=\lim_{h\to 0^-}\frac{|h|}{h}
    =\lim_{h\to 0^-}\frac{-h}{h}=-1 \\
    \lim_{h\to 0^+}\frac{|0+h|-|0|}{h}&=\lim_{h\to 0^+}\frac{|h|}{h}
    =\lim_{h\to 0^+}\frac{h}{h}=1
\end{align*}
thus the limit does not exist so \(f'(0)\) does not exist.
If a function is differentiable, then it is \textbf{smooth}, continuous, and
it has no vertical tangent lines.

\subsubsection{Higher Order Derivatives}
 If \(y=f(x)\) is a differentiable function and its derivative \(f'(x)\) is differentiable,
then the \textbf{second derivative} of \(f\) is
\[y''=f''(x)=\frac{d^2y}{dx^2}\]
We can interpret \(f''(x)\) as the slope of the curve \(y=f'(x)\) at the point
\((x,f'(x))\), which is the rate of change of the slope of the original curve \(y=f(x)\).
Let \(s=s(t)\) be a position function of an object with respect to time \(t\).
The velocity function \(v(t)\) of the object is \[v(t)=s'(t)=\frac{ds}{dt}\]
The instantaneous rate of change of velocity is the \textbf{acceleration}.
Thus the acceleration function \(a(t)\) is the derivative of the velocity function and
is therefore the second derivative of the position function.
\[a(t)=v'(t)=\frac{dv}{dt}=s''(t)=\frac{d^2s}{dt^2}\]
In general, the \(n\)th derivative of \(y=f(x)\) is
\[f^{(n)}(x)=\frac{d^ny}{dx^n}\]
\begin{problem}
    Find the first and the second derivatives of \(f(x)=x^3\).
\end{problem}
\begin{solution}
    We apply the \textbf{binomial theorem} by Newton
    \[(x+y)^n=\sum_{k=0}^n \binom{n}{k}x^{n-k}y^k\]
    For the first derivative we have
    \begin{align*}
        f'(x)&=\lim_{h\to 0}\frac{(x+h)^3-x^3}{h}=\lim_{h\to 0}
        \frac{x^3+3x^2h+3xh^2+h^3-x^3}{h}=\lim_{h\to 0}\frac{3x^2h+3xh^2+h^3}{h} \\
        &=\lim_{h\to 0}(3x^2+3xh+h^2)=3x^2
    \end{align*}
    For the second derivative we have
    \begin{align*}
        f''(x)&=\lim_{h\to 0}\frac{3(x+h)^2-3x^2}{h}
        =\lim_{h\to 0}\frac{3(x^2+2hx+h^2-x^2)}{h}=\lim_{h\to 0}\frac{6hx+3h^2}{h} \\
        &=\lim_{h\to 0}(6x+3h)=6x
    \end{align*}
    Therefore, the first and the second derivatives of \(f(x)=x^3\) are
    \begin{align*}
        &f'(x)=3x^2&f''(x)=6x&\qedhere
    \end{align*}
\end{solution}